---
title: "Delirium Prediction Model"
author: "Thomas Fisher"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(flextable)
library(ggplot2)
library(tidyverse)
library(here)
library(pROC)
library(patchwork)
library(xgboost)
library(glmnet)
library(dcurves)
library(tableone)
library(officer)
library(mgcv)
library(DiagrammeR)

# Null-coalescing helper
`%||%` <- function(a, b) if (!is.null(a)) a else b

# Safely load generated workspaces
load(here("data_interim", "model_cohorts.RData"))
load(here("data_interim", "results_descriptive.RData"))
load(here("data_interim", "sophisticated_results.RData"))
load(here("data_interim", "recalibration_results.RData"))
load(here("data_interim", "validation_results.RData"))

fmt_ci <- function(est, lo, hi, digits = 2, sep = " ") {
  f <- paste0("%.", digits, "f")
  dash <- if_else(lo < 0 | hi < 0, " to ", "–")
  paste0(sprintf(f, est), sep, "(", sprintf(f, lo), dash, sprintf(f, hi), ")")
}

# Fixed column widths (inches) so long cells and footnotes wrap instead of shrinking the font
fixed_widths <- function(ft, widths) {
  ft |> width(width = widths) |> set_table_properties(layout = "fixed")
}

# Define cohort_full dynamically if missing
if (!exists("cohort_full") && exists("cohort_raw_unimputed")) {
  cohort_full <- cohort_raw_unimputed
}

# Symmetrically load external comparisons if present, otherwise default to dynamic construction
if (file.exists(here("data_interim", "table_one_demographics.RData"))) {
  load(here("data_interim", "table_one_demographics.RData"))
}

# Alias modeling results
res <- sophisticated_results
rc <- recal_predictions

# CORE VARIABLES REFERENCE
CORE_VARS <- c("age_years", "urea_t0", "cog_pre_dementia_flag")

# JOURNAL TABLE STYLING HELPERS

journal_table <- function(ft, caption = "", footnote = NULL) {
  ft <- ft |>
    theme_booktabs() |>
    bold(part = "header") |>
    font(fontname = "Arial", part = "all") |>
    fontsize(size = 10, part = "all") |>
    fontsize(size = 9,  part = "footer") |>
    align(align = "center", part = "header") |>
    align(align = "left",   j = 1, part = "body") |>
    align(align = "center", j = -1, part = "body") |>
    set_caption(caption) |>
    padding(padding.top = 3, padding.bottom = 3, part = "all") |>
    autofit()
  if (!is.null(footnote))
    ft <- ft |> add_footer_lines(footnote)
  ft
}

# =============================================================================
# VARIABLE LABEL LOOKUP
# =============================================================================
var_labels <- c(
  age_years              = "Age, years",
  sex                    = "Sex, female",
  bmi                    = "Body mass index, kg/m²",
  asa_status             = "ASA physical status",
  fracture_type          = "Fracture type",
  cardiovascular         = "Cardiovascular disease",
  respiratory            = "Respiratory disease",
  cns                    = "CNS/neurological disease",
  urinary_tract          = "Urinary tract disease",
  cog_mmse_pre           = "MMSE score (pre-operative)",
  cog_pre_dementia_flag  = "Pre-existing dementia",
  temp_t0                = "Temperature, °C",
  sysbp_t0               = "Systolic blood pressure, mmHg",
  diabp_t0               = "Diastolic blood pressure, mmHg",
  hr_t0                  = "Heart rate, beats/min",
  spo2_t0                = "Oxygen saturation, %",
  hb_t0                  = "Haemoglobin, g/L",
  albumin_t0             = "Albumin, g/L",
  creatinine_t0          = "Creatinine, µmol/L",
  urea_t0                = "Urea, mmol/L",
  urea_creat_ratio_t0    = "Urea:creatinine ratio",
  na_t0                  = "Sodium, mmol/L",
  k_t0                   = "Potassium, mmol/L",
  cl_t0                  = "Chloride, mmol/L",
  alt_t0                 = "Alanine aminotransferase, U/L",
  ast_t0                 = "Aspartate aminotransferase, U/L",
  ph_t0                  = "Arterial pH",
  po2_kpa_t0             = "PaO₂, kPa",
  pco2_kpa_t0            = "PaCO₂, kPa",
  hco3_t0                = "Bicarbonate, mmol/L",
  anaes_tiva                 = "Anaesthesia: TIVA",
  anaes_volatile             = "Anaesthesia: volatile",
  anaes_regional             = "Anaesthesia: regional",
  duration_hrs               = "Surgical duration, hours",
  estimated_blood_loss       = "Estimated blood loss, mL",
  blood_transfusion          = "Intra-operative blood transfusion",
  intraop_hypotension        = "Intra-operative hypotension",
  hypotension_medicated      = "Hypotension requiring medication",
  intraop_propofol            = "Propofol (intra-operative)",
  intraop_etomidate           = "Etomidate (intra-operative)",
  intraop_sufentanil          = "Sufentanil (intra-operative)",
  intraop_remifentanil        = "Remifentanil (intra-operative)",
  intraop_fentanyl            = "Fentanyl (intra-operative)",
  intraop_morphine            = "Morphine (intra-operative)",
  intraop_butorphanol         = "Butorphanol (intra-operative)",
  intraop_tramadol            = "Tramadol (intra-operative)",
  intraop_hydromorphone       = "Hydromorphone (intra-operative)",
  intraop_oxycodone           = "Oxycodone (intra-operative)",
  intraop_ketorolac           = "Ketorolac (intra-operative)",
  intraop_nalbuphine          = "Nalbuphine (intra-operative)",
  intraop_lidocaine           = "Lidocaine (intra-operative)",
  intraop_ropivacaine         = "Ropivacaine (intra-operative)",
  intraop_levobupivacaine     = "Levobupivacaine (intra-operative)",
  intraop_bupivacaine         = "Bupivacaine (intra-operative)",
  intraop_chloroprocaine      = "Chloroprocaine (intra-operative)",
  intraop_atracurium          = "Atracurium (intra-operative)",
  intraop_rocuronium          = "Rocuronium (intra-operative)",
  intraop_vecuronium          = "Vecuronium (intra-operative)",
  intraop_atropine            = "Atropine (intra-operative)",
  intraop_adrenaline          = "Adrenaline (intra-operative)",
  intraop_digoxin             = "Digoxin (intra-operative)",
  intraop_anisodamine_nyxin   = "Anisodamine (intra-operative)",
  intraop_dexamethasone       = "Dexamethasone (intra-operative)",
  intraop_dexmedetomidine     = "Dexmedetomidine (intra-operative)",
  intraop_benzodiazepine      = "Benzodiazepine (intra-operative)",
  hydroxyethyl_starch = "Hydroxyethyl starch (colloid)",
  Anisodamine_nyxin = "Anisodamine"
)

label_vars <- function(x) {
  ifelse(x %in% names(var_labels), var_labels[x],
         str_replace_all(str_replace(x, "_t0$", ""), "_", " ") |> str_to_sentence())
}
```

# Main figures
## Figure 1: Patient Flow Diagram - manually drawn

## Supplementary Figure 4: Calibration of the models as developed (validation cohort)
```{r calibration_curves_main, fig.height=8, fig.width=12}

make_cal_plot_val <- function(obs, pred, title) {
  tibble(obs = as.numeric(as.character(obs)), pred = pred) |>
    filter(!is.na(obs) & !is.na(pred)) |>
    ggplot(aes(x = pred, y = obs)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
    geom_smooth(method = "loess", span = 0.75,
                colour = "steelblue", fill = "steelblue", alpha = 0.2) +
    geom_rug(alpha = 0.05, sides = "b") +
    coord_cartesian(xlim = c(0, 0.6), ylim = c(0, 0.6)) +
    theme_classic(base_family = "serif") +
    labs(title = title, x = "Predicted Probability", y = "Observed Proportion") +
    theme(
      plot.title = element_text(face = "bold", size = 10),
      axis.title = element_text(size = 10),
      axis.text = element_text(size = 9),
      plot.margin = margin(t = 5, r = 5, b = 5, l = 5)
    ) +
    # keep panels square to avoid vertical squashing when saved independently
    theme(aspect.ratio = 1)
}

y_val <- as.integer(res$obs_core_val)

fig2_plot <- (
  make_cal_plot_val(y_val, res$pred_l_core_val,     "Primary Model")  +
  make_cal_plot_val(y_val, res$pred_lasso_pre_val,  "Pre-op LASSO")   +
  make_cal_plot_val(y_val, res$pred_xgb_pre_val,    "Pre-op XGBoost")
) / (
  make_cal_plot_val(y_val, res$pred_lasso_post_val, "Intra-op LASSO")   +
  make_cal_plot_val(y_val, res$pred_xgb_post_val,   "Intra-op XGBoost") +
  plot_spacer()
)

fig2_plot

ggsave(here("outputs", "figures", "supplementary", "figureS4_calibration_as_developed.svg"), plot = fig2_plot,
       width = 180, height = 120, units = "mm", device = "svg")
ggsave(here("outputs", "figures", "supplementary", "figureS4_calibration_as_developed.png"), plot = fig2_plot,
       width = 180, height = 120, units = "mm", dpi = 300)
```

## Figure 3: Decision Curve Analysis (models as developed, and after cross-validated recalibration)
```{r DCA, fig.height=6, fig.width=10}
dca_colours <- c(
  "Primary Model"    = "#1b9e77",
  "Pre-op LASSO"     = "#d95f02",
  "Pre-op XGBoost"   = "#e7298a",
  "Intra-op LASSO"   = "#7570b3",
  "Intra-op XGBoost" = "#377eb8",
  "Treat all"        = "black",
  "Treat none"       = "gray70"
)

fig3_plot <- dca_curves |>
  mutate(
    Model = factor(Model, levels = names(dca_colours)),
    Predictions = factor(Predictions,
                         levels = c("Original", "Recalibrated (cross-validated)"),
                         labels = c("A. Models as developed", "B. After cross-validated recalibration"))
  ) |>
  ggplot(aes(x = threshold, y = net_benefit, colour = Model)) +
  geom_line(linewidth = 0.7) +
  facet_wrap(~ Predictions) +
  scale_colour_manual(values = dca_colours) +
  coord_cartesian(ylim = c(-0.02, 0.09), xlim = c(0, 0.25)) +
  theme_classic(base_family = "serif") +
  labs(
    x      = "Threshold Probability (Risk of Delirium)",
    y      = "Net Benefit",
    colour = NULL
  ) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.box = "vertical",
    legend.justification = "center",
    plot.title = element_text(face = "bold", size = 10),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 9),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9),
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 10, hjust = 0),
    plot.margin = margin(t = 5, r = 10, b = 30, l = 5)  # Increased bottom margin for legend
  )

fig3_plot

ggsave(here("outputs", "figures", "main", "figure3_DCA_validation.svg"), plot = fig3_plot,
       width = 220, height = 130, units = "mm", device = "svg")
ggsave(here("outputs", "figures", "main", "figure3_DCA_validation.png"), plot = fig3_plot,
       width = 220, height = 130, units = "mm", dpi = 300)
```

# Main tables
## Table 1: Cohort Overview
```{r cohort_overview}
build_table1_v4 <- function(df) {
  df_clean <- df |>
    mutate(
      cohort_assignment     = factor(cohort_assignment, levels = c("Derivation", "Validation")),
      delirium_7d           = as.integer(as.character(delirium_7d)),
      sex                   = as.integer(sex == "Female" | sex == 1 | sex == "1"),
      asa_status            = factor(case_when(
                                as.character(asa_status) == "1" ~ "ASA I",
                                as.character(asa_status) == "2" ~ "ASA II",
                                as.character(asa_status) == "3" ~ "ASA III",
                                as.character(asa_status) == "4" ~ "ASA IV",
                                TRUE ~ as.character(asa_status)
                              ), levels = c("ASA I", "ASA II", "ASA III", "ASA IV")),
      cog_pre_dementia_flag = as.integer(cog_pre_dementia_flag %in% c(1, TRUE, "TRUE")),
      fracture_type         = factor(fracture_type, levels = c("Intracapsular", "Extracapsular")),
      cardiovascular        = as.integer(as.integer(cardiovascular) == 1L),
      respiratory           = as.integer(as.integer(respiratory)    == 1L),
      cns                   = as.integer(as.integer(cns)            == 1L),
      urinary_tract         = as.integer(as.integer(urinary_tract)  == 1L)
    )

  # Six columns: each cohort in full, then split by delirium (del = NA means all patients)
  groups <- list(Deriv_All = c("Derivation", NA), Deriv_Del = c("Derivation", 1), Deriv_NoDel = c("Derivation", 0),
                 Val_All = c("Validation", NA),   Val_Del = c("Validation", 1),   Val_NoDel = c("Validation", 0))
  subset_of <- function(g) {
    d <- df_clean |> filter(cohort_assignment == g[1])
    if (!is.na(g[2])) d <- d |> filter(delirium_7d == as.integer(g[2]))
    d
  }
  mean_sd <- function(d, v) sprintf("%.1f (%.1f)", mean(d[[v]], na.rm = TRUE), sd(d[[v]], na.rm = TRUE))
  pct     <- function(d, v) { n <- sum(d[[v]], na.rm = TRUE); sprintf("%d (%.1f%%)", n, 100 * n / sum(!is.na(d[[v]]))) }
  cat_pct <- function(d, v, lvl) { n <- sum(d[[v]] == lvl, na.rm = TRUE); sprintf("%d (%.1f%%)", n, 100 * n / sum(!is.na(d[[v]]))) }

  row <- function(label, f, ...) {
    as_tibble(c(list(Variable = label), lapply(groups, function(g) f(subset_of(g), ...))))
  }
  hdr <- function(label) as_tibble(c(list(Variable = label), lapply(groups, function(g) "")))

  bind_rows(
    row("Age (years)",               mean_sd, "age_years"),
    row("Female sex",                pct,     "sex"),
    row("BMI (kg/m²)",          mean_sd, "bmi"),
    row("MMSE score",                mean_sd, "cog_mmse_pre"),
    row("Urea (mmol/L)",             mean_sd, "urea_t0"),
    row("Surgical duration (hours)", mean_sd, "duration_hrs"),
    hdr("ASA status"),
    purrr::map_dfr(levels(df_clean$asa_status), \(lvl) row(paste0("   ", lvl), cat_pct, "asa_status", lvl)),
    hdr("Fracture type"),
    purrr::map_dfr(levels(df_clean$fracture_type), \(lvl) row(paste0("   ", lvl), cat_pct, "fracture_type", lvl)),
    hdr("Comorbidities"),
    row("   Pre-existing dementia",  pct, "cog_pre_dementia_flag"),
    row("   Cardiovascular disease", pct, "cardiovascular"),
    row("   Respiratory disease",    pct, "respiratory"),
    row("   CNS disease",            pct, "cns"),
    row("   Urinary tract disease",  pct, "urinary_tract")
  )
}

t1 <- build_table1_v4(cohort_full)

# Dynamic counts
n_of <- function(cohort, del = NA) {
  d <- cohort_full |> filter(cohort_assignment == cohort)
  if (!is.na(del)) d <- d |> filter(as.integer(as.character(delirium_7d)) == del)
  nrow(d)
}

# Divider positions
header_rows <- which(t1$Variable %in% c(
  "Age (years)", "Female sex", "BMI (kg/m²)", "MMSE score", "Urea (mmol/L)", "Surgical duration (hours)",
  "ASA status", "Fracture type", "Comorbidities"
))
divider_after <- header_rows[-1] - 1

ft1 <- flextable(t1) |>
  set_header_labels(
    Variable    = "Characteristic",
    Deriv_All   = paste0("All\n(n = ", n_of("Derivation"), ")"),
    Deriv_Del   = paste0("Delirium\n(n = ", n_of("Derivation", 1), ")"),
    Deriv_NoDel = paste0("No delirium\n(n = ", n_of("Derivation", 0), ")"),
    Val_All     = paste0("All\n(n = ", n_of("Validation"), ")"),
    Val_Del     = paste0("Delirium\n(n = ", n_of("Validation", 1), ")"),
    Val_NoDel   = paste0("No delirium\n(n = ", n_of("Validation", 0), ")")
  ) |>
  add_header_row(values = c("", "Derivation cohort", "Validation cohort"), colwidths = c(1, 3, 3)) |>
  theme_booktabs() |>
  align(align = "center", part = "header") |>
  align(align = "left",   j = 1,   part = "body") |>
  align(align = "center", j = 2:7, part = "body") |>
  bold(part = "header") |>
  bold(j = c(2, 5), part = "body") |>
  bold(i = which(t1$Variable %in% c("ASA status", "Fracture type", "Comorbidities")),
       j = 1, part = "body") |>
  hline(i = divider_after,
        border = officer::fp_border(color = "#CCCCCC", width = 0.5),
        part = "body") |>
  add_footer_lines(values = c(
    paste0("Continuous variables reported as mean (SD) of observed values; categorical and binary variables as ",
           "n (%). Missing data for each variable are shown in Supplementary Table S3.")
  )) |>
  italic(part = "footer") |>
  font(fontname = "Arial", part = "all") |>
  fontsize(size = 8, part = "all") |>
  set_caption("Table 1. Demographic and clinical characteristics of the derivation and validation cohorts, overall and by delirium occurrence") |>
  fixed_widths(c(1.62, rep(0.78, 6)))

# Save Lancet-friendly versions
flextable::save_as_docx(ft1, path = here("outputs", "tables", "main", "table1_cohort_overview.docx"))
ft1
```

## Table 2: Model Coefficients
```{r coefs_pre}
prep_df <- function(df, model_name) {
  df |>
    mutate(
      Predictor = label_vars(Predictor),
      !!model_name := sprintf("%.3f
%.2f (%.2f–%.2f)", Beta, OR, LCI, UCI),
      Beta_val = abs(Beta)
    ) |>
    select(Predictor, !!model_name, Beta_val)
}

combined_coefs <- prep_df(res$coefs_core, "Primary model") |>
  full_join(prep_df(res$coefs_pre, "Pre-op LASSO"), by = "Predictor", suffix = c("_core", "_pre")) |>
  full_join(prep_df(res$coefs_post, "Intra-op LASSO"), by = "Predictor") |>
  mutate(max_beta = pmax(Beta_val_core, Beta_val_pre, Beta_val, na.rm = TRUE)) |>
  arrange(desc(max_beta)) |>
  select(Predictor, `Primary model`, `Pre-op LASSO`, `Intra-op LASSO`)

b0_core  <- res$core_intercept %||% -3.124
b0_pre   <- res$pre_intercept  %||% -3.245
b0_post  <- res$post_intercept %||% -3.512

intercept_row <- tibble(
  Predictor = "Intercept",
  `Primary model`  = sprintf("%.3f", b0_core),
  `Pre-op LASSO`   = sprintf("%.3f", b0_pre),
  `Intra-op LASSO` = sprintf("%.3f", b0_post)
)

combined_coefs <- bind_rows(intercept_row, combined_coefs)

ft2 <- combined_coefs |>
  flextable() |>
  set_header_labels(Predictor = "Predictor") |>
  add_header_row(
    values = c("", "β coefficient; odds ratio (95% CI)"),
    colwidths = c(1, 3)
  ) |>
  theme_booktabs() |>
  colformat_char(na_str = "—") |>
  align(align = "center", j = 2:4, part = "all") |>
  add_footer_lines(paste0(
    "Each cell shows the log-odds β coefficient (first line) and the corresponding odds ratio with 95% CI (second line). ",
    "The linear predictor z (P = 1/[1+exp(−z)]) is the intercept plus the sum of each retained predictor's β ",
    "multiplied by its value; intercept rows show the log-odds constant only. For the LASSO models, coefficients and CIs ",
    "come from an unpenalised refit of the stability-selected predictors and do not account for the selection step, so ",
    "the CIs are too narrow. Predictors are ordered by descending absolute β; coefficients rounded to three decimal places."
  )) |>
  set_caption("Table 2. Predictors included in the final prediction models") |>
  fixed_widths(c(2.1, 1.4, 1.4, 1.4))

# Save Lancet-friendly versions
flextable::save_as_docx(ft2, path = here("outputs", "tables", "main", "table2_model_coefficients.docx"))
ft2
```

## Table 3: Model performance
```{r model_performance_comparison}
model_order <- c("Primary Model", "Pre-op LASSO", "Pre-op XGBoost", "Intra-op LASSO", "Intra-op XGBoost")

internal_val <- res$internal_validation |>
  mutate(Model = if_else(Model == "Core Model", "Primary Model", Model))

# XGBoost predictions are already compressed in the derivation data (apparent slope well above 1),
# so a bootstrap-corrected slope is not interpretable for them; show the apparent slope instead.
deriv_slope <- function(obs, p) {
  lp <- qlogis(pmin(pmax(p, 1e-4), 1 - 1e-4))
  unname(coef(glm(as.integer(as.character(obs)) ~ lp, family = "binomial"))[2])
}
xgb_apparent_slope <- c(
  "Pre-op XGBoost"   = deriv_slope(res$obs_pre,  res$pred_xgb_pre),
  "Intra-op XGBoost" = deriv_slope(res$obs_post, res$pred_xgb_post)
)

table3_data <- internal_val |>
  select(Model, Apparent_AUC, Optimism_Corrected_AUC, Corrected_Slope) |>
  left_join(val_metrics, by = "Model") |>
  mutate(Model = factor(Model, levels = model_order)) |>
  arrange(Model) |>
  transmute(
    Model             = as.character(Model),
    `Apparent AUROC`  = sprintf("%.3f", Apparent_AUC),
    `Corrected AUROC` = sprintf("%.3f", Optimism_Corrected_AUC),
    `Corrected slope` = if_else(Model %in% names(xgb_apparent_slope),
                                sprintf("%.2f*", xgb_apparent_slope[Model]),
                                sprintf("%.2f", Corrected_Slope)),
    `AUROC (95% CI)` = fmt_ci(AUC, AUC_LCI, AUC_UCI, 3, "
"),
    `CITL (95% CI)`  = fmt_ci(CITL, CITL_LCI, CITL_UCI, 2, "
"),
    `Slope (95% CI)` = fmt_ci(Slope, Slope_LCI, Slope_UCI, 2, "
")
  )

ft3 <- table3_data |>
  flextable() |>
  theme_booktabs() |>
  add_header_row(
    values = c("", "Internal validation (derivation cohort)", "Validation cohort (models as developed)"),
    colwidths = c(1, 3, 3)
  ) |>
  bold(part = "header") |>
  font(fontname = "Arial", part = "all") |>
  fontsize(size = 8, part = "all") |>
  align(align = "center", part = "header") |>
  align(align = "center", j = 2:7, part = "body") |>
  valign(valign = "center", part = "body") |>
  hline(i = 1, border = officer::fp_border(width = 1.2, color = "black")) |>
  set_caption("Table 3. Model discrimination and calibration in the derivation and validation cohorts") |>
  add_footer_lines(paste0(
    "AUROC = area under the receiver operating characteristic curve. CITL = calibration-in-the-large ",
    "(intercept of a logistic model with the model's linear predictor as an offset; 0 is ideal, positive values ",
    "indicate under-prediction). Slope = calibration slope (1 is ideal). Optimism correction used Harrell's bootstrap ",
    "with every modelling step repeated in each replicate: 1000 replicates for the primary model, 200 for the LASSO ",
    "models (including stability selection) and 1000 for XGBoost (including tuning of the number of boosting rounds). ",
    "The corrected slope is the bootstrap estimate of the calibration slope expected in new data. ",
    "*Apparent calibration slope in the derivation cohort: XGBoost predictions were compressed towards the average ",
    "risk even in the data used to fit them, so a bootstrap-corrected slope is not interpretable. ",
    "Validation metrics use each model's original predictions without recalibration; 95% CIs for AUROC by ",
    "DeLong's method and Wald 95% CIs for calibration measures."
  )) |>
  fontsize(size = 7.5, part = "footer") |>
  fixed_widths(c(1.1, 0.75, 0.75, 0.75, 1.05, 0.95, 0.95))

# Save Lancet-friendly versions
flextable::save_as_docx(ft3, path = here("outputs", "tables", "main", "table3_model_performance.docx"))
ft3
```

# Supplementary Figures
## Supplementary Figure 1: Measured values vs delirium risk

```{r lab_gam_plots, fig.height=8, fig.width=12}
lab_src <- if (exists("cohort_raw_descriptive")) cohort_raw_descriptive else cohort_full

lab_cols_present <- names(lab_src)[
  str_detect(names(lab_src), "_t0$") &
  !str_detect(names(lab_src), "delirium|pain|map|shock|urea_creat")
]

if (length(lab_cols_present) > 0 && "delirium_7d" %in% names(lab_src)) {
  lab_long_rmd <- lab_src |>
    select(delirium_7d, all_of(lab_cols_present)) |>
    mutate(delirium_7d = as.numeric(as.character(delirium_7d))) |>
    pivot_longer(-delirium_7d, names_to = "lab", values_to = "value") |>
    filter(!is.na(value), !is.na(delirium_7d)) |>
    mutate(lab_label = label_vars(str_remove(lab, "_t0$")))

  ggplot(lab_long_rmd, aes(x = value, y = delirium_7d)) +
    geom_jitter(height = 0.05, alpha = 0.2, size = 0.5) +
    geom_smooth(method = "glm", method.args = list(family = "binomial"),
                color = "red", linetype = "dashed", se = FALSE) +
    geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"),
                color = "blue", se = FALSE) +
    coord_cartesian(ylim = c(0, 1)) +
    facet_wrap(~lab_label, scales = "free_x") +
    theme_minimal() +
    labs(title    = "Lab Values vs. Delirium Risk",
         subtitle = "Red = Linear (logistic) | Blue = Non-linear (GAM)",
         y = "Delirium Probability", x = "Value")
} else {
  message("Lab columns or outcome not available — skipping lab plots.")
}

```

## Supplementary Figure 2: GLM and LASSO model equations (uncalibrated)
```{r glm_core_stacked, results='asis'}
b0 <- res$core_intercept %||% -3.124

coefs_core_active <- res$coefs_core |> 
  filter(Predictor != "(Intercept)")

terms_core_stacked <- purrr::map2_chr(
  coefs_core_active$Beta, 
  label_vars(coefs_core_active$Predictor), 
  function(beta, label) {
    sign_str <- if (beta >= 0) "+ " else "- "
    sprintf("  & %s %.3f \\times \\text{%s} \\\\", sign_str, abs(beta), label)
  }
)
cat("Primary Model:")
cat("$$P(\\text{Delirium}) = \\frac{1}{1 + \\exp(-z)}$$\n\n")

cat("Where $z$ is resolved as:\n\n")
cat("$$\\begin{aligned}\n")
cat(sprintf("z = & %.3f \\\\\n", b0))
cat(paste(terms_core_stacked, collapse = "\n"))
cat("\\end{aligned}$$\n\n")
```

```{r lasso_pre_stacked, results='asis'}
b0_pre   <- res$pre_intercept %||% -3.245

coefs_pre_active <- res$coefs_pre |> 
  filter(Beta != 0, Predictor != "(Intercept)")

terms_pre_stacked <- purrr::map2_chr(
  coefs_pre_active$Beta, 
  label_vars(coefs_pre_active$Predictor), 
  function(beta, label) {
    sign_str <- if (beta >= 0) "+ " else "- "
    sprintf("  & %s %.3f \\times \\text{%s} \\\\", sign_str, abs(beta), label)
  }
)
cat("Pre-operative LASSO Model:")
cat("$$P(\\text{Delirium}) = \\frac{1}{1 + \\exp(-z)}$$\n\n")

cat("Where $z$ is resolved as:\n\n")
cat("$$\\begin{aligned}\n")
cat(sprintf("z = & %.3f \\\\\n", b0_pre))
cat(paste(terms_pre_stacked, collapse = "\n"))
cat("\\end{aligned}$$\n\n")
```
```{r lasso_post_stacked, results='asis'}
b0_post  <- res$post_intercept %||% -3.512

coefs_post_active <- res$coefs_post |> 
  filter(Beta != 0, Predictor != "(Intercept)")

terms_post_stacked <- purrr::map2_chr(
  coefs_post_active$Beta, 
  label_vars(coefs_post_active$Predictor), 
  function(beta, label) {
    sign_str <- if (beta >= 0) "+ " else "- "
    sprintf("  & %s %.3f \\times \\text{%s} \\\\", sign_str, abs(beta), label)
  }
)
cat("Intraoperative LASSO Model:")

cat("$$P(\\text{Delirium}) = \\frac{1}{1 + \\exp(-z)}$$\n\n")

cat("Where $z$ is resolved as:\n\n")
cat("$$\\begin{aligned}\n")
cat(sprintf("z = & %.3f \\\\\n", b0_post))
cat(paste(terms_post_stacked, collapse = "\n"))
cat("\\end{aligned}$$\n\n")
```

## Supplementary Figure 3: XGBoost Feature Importance
```{r xgb_importance, fig.height=6, fig.width=12}
relabel_importance <- function(imp) {
  imp |>
    mutate(Feature = label_vars(
      str_remove(Feature, "(TRUE|FALSE|[.]Missing|[.]Measured)$")
    ))
}

imp_pre  <- res$xgb_importance_pre
imp_post <- res$xgb_importance_post

p_imp_pre <- xgb.ggplot.importance(relabel_importance(imp_pre), top_n = 12) +
  theme_classic(base_family = "serif") +
  labs(title = "A. Pre-operative XGBoost", x = NULL, y = "Information Gain") +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 10),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 9)
  )

p_imp_post <- xgb.ggplot.importance(relabel_importance(imp_post), top_n = 12) +
  theme_classic(base_family = "serif") +
  labs(title = "B. Intraoperative XGBoost", x = NULL, y = "Information Gain") +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 10),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 9)
  )

figS3_plot <- p_imp_pre + p_imp_post +
  plot_annotation(
    title    = "XGBoost Feature Importance (Top 12 by Information Gain)",
    subtitle = "Features ordered by mean gain across trees",
    theme = theme(plot.title = element_text(family = "serif", face = "bold", size = 10), plot.subtitle = element_text(family = "serif", size = 9))
  )

figS3_plot

ggsave(here("outputs", "figures", "supplementary", "figureS3_xgb_importance.svg"), plot = figS3_plot,
       width = 180, height = 90, units = "mm", device = "svg")
ggsave(here("outputs", "figures", "supplementary", "figureS3_xgb_importance.png"), plot = figS3_plot,
       width = 180, height = 90, units = "mm", dpi = 300)
```

## Figure 2: Calibration after cross-validated recalibration (validation cohort)
```{r calibration_curves_raw_supp, fig.height=8, fig.width=12}
# Calibration Curves (Validation Cohort LOESS Fits - Uncalibrated / Raw)
make_cal_plot_val_raw <- function(obs, pred, title) {
  tibble(obs = obs, pred = pred) |>
    filter(!is.na(obs) & !is.na(pred)) |>
    ggplot(aes(x = pred, y = obs)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
    geom_smooth(method = "loess", span = 0.75,
                colour = "steelblue", fill = "steelblue", alpha = 0.2) +
    geom_rug(alpha = 0.05, sides = "b") +
    coord_cartesian(xlim = c(0, 0.6), ylim = c(0, 0.6)) +
    theme_classic(base_family = "serif") +
    labs(title = title, x = "Predicted Probability", y = "Observed Proportion") +
    theme(
      plot.title = element_text(face = "bold", size = 10),
      axis.title = element_text(size = 10),
      axis.text = element_text(size = 9)
    )
}

figS4_plot <- (
  make_cal_plot_val_raw(y_val, preds_recal_cv[["Primary Model"]],    "Primary Model")  +
  make_cal_plot_val_raw(y_val, preds_recal_cv[["Pre-op LASSO"]],     "Pre-op LASSO")   +
  make_cal_plot_val_raw(y_val, preds_recal_cv[["Pre-op XGBoost"]],   "Pre-op XGBoost")
) / (
  make_cal_plot_val_raw(y_val, preds_recal_cv[["Intra-op LASSO"]],   "Intra-op LASSO")   +
  make_cal_plot_val_raw(y_val, preds_recal_cv[["Intra-op XGBoost"]], "Intra-op XGBoost") +
  plot_spacer()
)

figS4_plot

ggsave(here("outputs", "figures", "main", "figure2_calibration_recalibrated.svg"), plot = figS4_plot,
       width = 180, height = 120, units = "mm", device = "svg")
ggsave(here("outputs", "figures", "main", "figure2_calibration_recalibrated.png"), plot = figS4_plot,
       width = 180, height = 120, units = "mm", dpi = 300)
```

## Supplementary tables
# Supplementary Table 1: Intraoperative drug exclusion summary
```{r intraop-prevalence-table, echo=FALSE}
library(tidyverse)
library(flextable)

all_intraop_vars <- names(cohort_postop_df)[grep("^intraop_|^intra_operative_", names(cohort_postop_df))]

drug_prev_df <- cohort_postop_df |>
  filter(cohort_assignment == "Derivation") |>
  summarise(across(all_of(all_intraop_vars), ~ mean(.x != 0, na.rm = TRUE))) |>
  pivot_longer(cols = everything(), names_to = "Raw_Variable", values_to = "Prevalence") |>
  filter(!str_detect(Raw_Variable, "(?i)hypotension|duration|loss|transfusion|blood|type|anaesthesia")) |>
  mutate(
    Predictor = str_to_title(str_replace_all(Raw_Variable, "^intraop_|^intra_operative_", "")),
    Prevalence_Pct = Prevalence * 100,
    Status = ifelse(Prevalence >= 0.05, "Retained (\u22655%)", "Filtered Out (<5%)")
  ) |>
  arrange(desc(Prevalence))

prev_table <- flextable(drug_prev_df |> select(Predictor, Prevalence_Pct, Status)) |>
  set_header_labels(
    Predictor = "Intraoperative Agent/Class",
    Prevalence_Pct = "Cohort Prevalence (%)",
    Status = "Retained?"
  ) |>
  colformat_double(j = "Prevalence_Pct", digits = 2) |>
  autofit() |>
  theme_booktabs() |>
  color(i = ~ Prevalence_Pct < 5, color = "darkorange") |>
  italic(i = ~ Prevalence_Pct < 5)

flextable::save_as_docx(prev_table, path = here("outputs", "tables", "supplementary", "tableS1_intraop_prevalence.docx"))
prev_table
```

## Supplementary Table 2: Candidate Predictor Summary
```{r candidate_predictors}
pred_table <- tibble::tribble(
  ~Predictor,                         ~`Pre-op models`, ~`Intra-op models`, ~`Primary model`,
  "Age, years",                        "✓", "✓", "✓",
  "Sex",                               "✓", "✓", "",
  "BMI, kg/m²",                        "✓", "✓", "",
  "ASA physical status",               "✓", "✓", "",
  "Fracture type",                     "✓", "✓", "",
  "Cardiovascular disease",            "✓", "✓", "",
  "Respiratory disease",               "✓", "✓", "",
  "CNS/neurological disease",          "✓", "✓", "",
  "Urinary tract disease",             "✓", "✓", "",
  "Pre-existing dementia",             "✓", "✓", "✓",
  "MMSE score",                        "✓", "✓", "",
  "Education level",                   "✓", "✓", "",
  "Temperature, °C",                   "✓", "✓", "",
  "Systolic blood pressure, mmHg",     "✓", "✓", "",
  "Diastolic blood pressure, mmHg",    "✓", "✓", "",
  "Heart rate, beats/min",             "✓", "✓", "",
  "Haemoglobin, g/L",                  "✓", "✓", "",
  "Albumin, g/L",                      "✓", "✓", "",
  "Creatinine, µmol/L",               "✓", "✓", "",
  "Urea, mmol/L",                      "✓", "✓", "✓",
  "Urea:creatinine ratio",             "✓", "✓", "",
  "Sodium, mmol/L",                    "✓", "✓", "",
  "Potassium, mmol/L",                 "✓", "✓", "",
  "Chloride, mmol/L",                  "✓", "✓", "",
  "ALT, U/L",                          "✓", "✓", "",
  "AST, U/L",                          "✓", "✓", "",
  "Oxygen saturation, %",              "✓", "✓", "",
  "Anaesthesia type (TIVA/Volatile/Regional)", "", "✓", "",
  "Intra-operative hypotension",       "",  "✓", "",
  "Hypotension requiring medication",  "",  "✓", "",
  "Estimated blood loss, mL",          "",  "✓", "",
  "Surgical duration, hours",          "",  "✓", "",
  "Blood transfusion",                 "",  "✓", "",
  "Intra-operative agents (Propofol, Sufentanil, Remifentanil, Fentanyl, Lidocaine, Ropivacaine, Bupivacaine, Atracurium, Rocuronium, Etomidate)", "", "✓", ""
)

pred_table |>
  flextable() |>
  theme_booktabs() |>
  bold(part = "header") |>
  align(align = "center", j = 2:4, part = "all") |>
  bg(i = ~ `Pre-op models` == "" & `Intra-op models` == "", j = 1:4, bg = "#f8f8f8") |>
  set_caption("Table S2. Candidate predictors made available to each model by timing of availability") |>
  add_footer_lines(paste0(
    "✓ = predictor made available to this model. Pre-op and intra-op models = the LASSO and XGBoost models ",
    "for each time point, which shared the same candidate predictors. Missing continuous values were imputed ",
    "with derivation-cohort medians. Arterial blood gas values (pH, PaO₂, PaCO₂, bicarbonate) were not offered to the models because they were missing in over half of patients (Table S3)."
  )) |>
  fontsize(size = 9, part = "all") |>
  fixed_widths(c(3.6, 0.9, 0.9, 0.9)) |>
  (\(x) {
    flextable::save_as_docx(x, path = here("outputs", "tables", "supplementary", "tableS2_candidate_predictors.docx"))
    x
  })()
```

## Supplementary Table 3: Summary of missingness
```{r cohort-overview-comprehensive}
new_drug_vars <- names(cohort_full) %>% 
  str_subset("^intraop_") %>% 
  setdiff(c("intraop_hypotension"))

all_model_vars <- unique(c(
  t0_vars, 
  lab_vars, 
  abg_vars,
  "map_t0", 
  "urea_creat_ratio_t0", 
  "blood_transfusion",
  new_drug_vars 
)) |>
  setdiff(c("blood_transfusion_units", "total_transfusion", "estimated_blood_loss"))

get_stats_by_cohort <- function(v_name, df) {
  v_deriv <- df[[v_name]][df$cohort_assignment == "Derivation"]
  v_val   <- df[[v_name]][df$cohort_assignment == "Validation"]
  
  helper <- function(v) {
    n_total  <- length(v)
    n_miss   <- sum(is.na(v))
    miss_pct <- (n_miss / n_total) * 100
    v_clean  <- v[!is.na(v)]
    n_valid  <- length(v_clean)
    
    if (n_valid == 0) {
      return(list(missing = "0 (0.0%)", summary = "—", range = "—", miss_pct = 0))
    }
    
    is_cat <- is.factor(v) || is.character(v) || is.logical(v) ||
              (is.numeric(v) && all(v_clean %in% c(0, 1, 1L, 0L)))
              
    if (is_cat) {
      positive_labels <- c("1", "TRUE", "Female", "Yes")
      if (any(as.character(v_clean) %in% positive_labels)) {
        n_yes <- sum(v_clean == 1 | v_clean == TRUE | v_clean == "Female" | v_clean == "Yes", na.rm = TRUE)
        stats <- paste0(n_yes, " (", round(n_yes / n_valid * 100, 1), "%)")
      } else {
        # not a recognised binary-positive label (e.g. fracture type) -- report
        # the modal category by name rather than a meaningless 0 count
        tab      <- sort(table(v_clean), decreasing = TRUE)
        top_level <- names(tab)[1]
        stats    <- paste0(top_level, ", ", tab[1], " (", round(tab[1] / n_valid * 100, 1), "%)")
      }
      rng   <- "—"
    } else {
      v_num <- as.numeric(v_clean)
      stats <- paste0(round(mean(v_num), 1), " (", round(sd(v_num), 1), ")")
      rng   <- paste0(round(min(v_num), 1), " \u2013 ", round(max(v_num), 1))
    }
    
    list(
      missing = sprintf("%d (%.1f%%)", n_miss, miss_pct),
      summary = stats,
      range = rng,
      miss_pct = miss_pct
    )
  }
  
  d_stats <- helper(v_deriv)
  v_stats <- helper(v_val)
  
  tibble(
    Variable       = v_name,
    Deriv_Missing  = d_stats$missing,
    Deriv_Summary  = d_stats$summary,
    Deriv_Range    = d_stats$range,
    Val_Missing    = v_stats$missing,
    Val_Summary    = v_stats$summary,
    Val_Range      = v_stats$range,
    max_miss_pct   = max(d_stats$miss_pct, v_stats$miss_pct)
  )
}

src_df <- if (exists("cohort_cleaned_preimp")) cohort_cleaned_preimp else cohort_full

overview_df <- all_model_vars |>
  intersect(names(src_df)) |>
  map_df(~get_stats_by_cohort(.x, src_df)) |>
  mutate(
    Variable_Clean = str_replace(Variable, "_t0$", "") |>
                     str_replace_all("_", " ") |>
                     str_to_title(),
    Label = case_match(
      Variable,
      "age_years"             ~ "Age at admission, years",
      "sex"                   ~ "Sex, female",
      "bmi"                   ~ "BMI, kg/m\u00b2",
      "asa_status"            ~ "ASA physical status",
      "fracture_type"         ~ "Fracture type",
      "cardiovascular"        ~ "Cardiovascular disease",
      "respiratory"           ~ "Respiratory disease",
      "cns"                   ~ "CNS/neurological disease",
      "urinary_tract"         ~ "Urinary tract disease",
      "cog_mmse_pre"          ~ "MMSE score (pre-operative)",
      "cog_pre_dementia_flag" ~ "Pre-existing dementia",
      "temp_t0"               ~ "Temperature, \u00b0C",
      "sysbp_t0"              ~ "Systolic BP, mmHg",
      "diabp_t0"              ~ "Diastolic BP, mmHg",
      "map_t0"                ~ "Mean arterial pressure, mmHg",
      "hr_t0"                 ~ "Heart rate, beats/min",
      "spo2_t0"               ~ "SpO\u2082, %",
      "hb_t0"                 ~ "Haemoglobin, g/L",
      "albumin_t0"            ~ "Albumin, g/L",
      "creatinine_t0"         ~ "Creatinine, \u00b5mol/L",
      "urea_t0"               ~ "Urea, mmol/L",
      "urea_creat_ratio_t0"   ~ "Urea:creatinine ratio",
      "na_t0"                 ~ "Sodium, mmol/L",
      "k_t0"                  ~ "Potassium, mmol/L",
      "cl_t0"                 ~ "Chloride, mmol/L",
      "alt_t0"                ~ "Alanine aminotransferase, U/L",
      "ast_t0"                ~ "Aspartate aminotransferase, U/L",
      "ph_t0"                 ~ "Arterial pH",
      "po2_kpa_t0"            ~ "PaO\u2082, kPa",
      "pco2_kpa_t0"           ~ "PaCO\u2082, kPa",
      "hco3_t0"               ~ "Bicarbonate, mmol/L",
      "fluphenazine"              ~ "Fluphenazine",
      "benzodiazepine_sedatives"  ~ "Benzodiazepine/sedatives",
      "opioid_analgesics"         ~ "Opioid analgesics",
      "nonsteroidal_analgesics"   ~ "NSAIDs",
      "other_psychotropic_drugs"  ~ "Other psychotropic drugs",
      "anaes_tiva"                ~ "Anaesthesia: TIVA",
      "anaes_volatile"            ~ "Anaesthesia: volatile",
      "anaes_regional"            ~ "Anaesthesia: regional",
      .default = Variable_Clean
    )
  )

# Get dynamic counts for the subheaders
n_deriv <- sum(cohort_full$cohort_assignment == "Derivation")
n_val   <- sum(cohort_full$cohort_assignment == "Validation")

tableS3 <- flextable(overview_df,
          col_keys = c("Label", 
                       "Deriv_Missing", "Deriv_Summary", "Deriv_Range", 
                       "Val_Missing", "Val_Summary", "Val_Range")) |>
  set_header_labels(
    Label          = "Variable",
    Deriv_Missing  = "Missing, n (%)",
    Deriv_Summary  = "Mean (SD) or n (%)",
    Deriv_Range    = "Observed range",
    Val_Missing    = "Missing, n (%)",
    Val_Summary    = "Mean (SD) or n (%)",
    Val_Range      = "Observed range"
  ) |>
  add_header_row(
    values = c("", 
               paste0("Derivation Cohort (n = ", n_deriv, ")"), 
               paste0("Validation Cohort (n = ", n_val, ")")),
    colwidths = c(1, 3, 3)
  ) |>
  color(i = ~max_miss_pct > 5,  j = c("Deriv_Missing", "Val_Missing"), color = "firebrick") |>
  color(i = ~max_miss_pct > 20, j = c("Deriv_Missing", "Val_Missing"), color = "darkred")   |>
  bold( i = ~max_miss_pct > 5,  j = c("Deriv_Missing", "Val_Missing"))  |>
  theme_booktabs() |>
  set_caption("Table S3. Data completeness and descriptive summary of all candidate predictors split by cohort assignment") |>
  add_footer_lines(paste0(
    "Values are mean (SD) for continuous variables or n (%) for binary/categorical variables. ",
    "Missing values flagged in red (>5%) or dark red (>20%). Arterial blood gas values were not used as candidate predictors."
  )) |>
  italic(part = "footer") |>
  fontsize(size = 9, part = "footer")

# Save supplementary table
flextable::save_as_docx(tableS3, path = here("outputs", "tables", "supplementary", "tableS3_data_completeness.docx"))
tableS3
```

## Supplementary Table 4: Delirium Incidence by Centre
```{r incidence_by_centre}
ft_centre <- cohort_raw_unimputed |>
  group_by(Centre = centre) |>
  summarise(
    n = n(),
    `Delirium, n` = sum(delirium_7d == 1, na.rm = TRUE),
    `Delirium incidence, %` = sprintf("%.1f", 100 * mean(delirium_7d == 1, na.rm = TRUE)),
    `Missing outcome, n` = sum(is.na(delirium_7d)),
    .groups = "drop"
  ) |>
  arrange(desc(n)) |>
  flextable() |>
  journal_table(
    caption  = "Table S4. Delirium incidence by recruiting centre",
    footnote = "Incidence is the percentage of patients with delirium within 7 days of surgery."
  ) |>
  fixed_widths(c(3.1, 0.6, 0.8, 0.9, 0.9))

# Save to your supplementary tables folder
flextable::save_as_docx(ft_centre, path = here("outputs", "tables", "supplementary", "tableS4_incidence_by_centre.docx"))
ft_centre
```

# Supplementary Table 5: Threshold metrics and net benefit
```{r clinical_threshold_metrics}
s5_data <- nb_table |>
  filter(Model != "Treat all") |>
  mutate(
    Model = factor(Model, levels = model_order),
    Predictions = if_else(Predictions == "Original", "Models as developed", "After cross-validated recalibration"),
    Predictions = factor(Predictions, levels = c("Models as developed", "After cross-validated recalibration"))
  ) |>
  arrange(Predictions, threshold, Model) |>
  left_join(
    nb_table |> filter(Model == "Treat all") |> select(threshold, NB_all = NB),
    by = "threshold"
  ) |>
  transmute(
    Predictions = as.character(Predictions),
    Threshold   = sprintf("%.0f%%", threshold * 100),
    Model       = as.character(Model),
    `Above threshold, %` = sprintf("%.1f", Flagged_Pct * 100),
    Sensitivity = sprintf("%.1f%%", Sensitivity * 100),
    Specificity = sprintf("%.1f%%", Specificity * 100),
    PPV         = if_else(is.na(PPV), "—", sprintf("%.1f%%", PPV * 100)),
    NPV         = sprintf("%.1f%%", NPV * 100),
    `Net benefit (95% CI)` = fmt_ci(NB, NB_LCI, NB_UCI, 3),
    `Treat all` = sprintf("%.3f", NB_all)
  )

ft5 <- s5_data |>
  flextable() |>
  theme_booktabs() |>
  bold(part = "header") |>
  font(fontname = "Arial", part = "all") |>
  fontsize(size = 8, part = "all") |>
  align(align = "center", part = "header") |>
  align(align = "left",   j = 1:3, part = "body") |>
  align(align = "center", j = 4:10, part = "body") |>
  merge_v(j = c(1, 2, 10)) |>
  valign(j = c(1, 2, 10), valign = "top") |>
  hline(i = seq(5, nrow(s5_data) - 1, by = 5), border = officer::fp_border(color = "#CCCCCC", width = 0.5)) |>
  set_caption("Table S5. Classification performance and net benefit at clinically relevant risk thresholds in the validation cohort") |>
  add_footer_lines(paste0(
    "Above threshold: patients with predicted risk at or above the threshold. Net benefit 95% CIs by bootstrap. ",
    "Recalibration was logistic, cross-validated within the validation cohort. As developed, XGBoost predictions rarely ",
    "reached 10% (maximum 11% and 9%)."
  )) |>
  italic(part = "footer") |>
  fontsize(size = 7.5, part = "footer") |>
  fixed_widths(c(1.35, 0.65, 1.1, 0.7, 0.8, 0.8, 0.65, 0.65, 1.45, 0.6))

flextable::save_as_docx(
  ft5,
  path = here("outputs", "tables", "supplementary", "tableS5_clinical_thresholds.docx"),
  pr_section = officer::prop_section(page_size = officer::page_size(orient = "landscape"))
)
ft5
```

# Supplementary Table 6: Detailed validation metrics
```{r validation_detail}
y_val_s6 <- as.integer(res$obs_core_val)
brier_null <- mean(y_val_s6) * (1 - mean(y_val_s6))   # Brier of predicting the cohort incidence for everyone

s6_data <- val_metrics |>
  left_join(auc_diff, by = "Model") |>
  mutate(Model = factor(Model, levels = model_order)) |>
  arrange(Model) |>
  transmute(
    Model = as.character(Model),
    `Brier (95% CI)` = fmt_ci(Brier, Brier_LCI, Brier_UCI, 3, "
"),
    `ICI (95% CI)`   = fmt_ci(ICI, ICI_LCI, ICI_UCI, 3, "
"),
    `AUROC difference from primary model (95% CI)` = if_else(is.na(Diff), "—", fmt_ci(Diff, LCI, UCI, 3, "
"))
  )

ft6 <- s6_data |>
  flextable() |>
  theme_booktabs() |>
  bold(part = "header") |>
  font(fontname = "Arial", part = "all") |>
  fontsize(size = 8, part = "all") |>
  align(align = "center", part = "header") |>
  align(align = "center", j = 2:4, part = "body") |>
  set_caption("Table S6. Brier score and integrated calibration index in the validation cohort, and AUROC difference from the primary model") |>
  add_footer_lines("ICI, integrated calibration index. Confidence intervals by bootstrap.") |>
  fontsize(size = 7.5, part = "footer") |>
  fixed_widths(c(1.5, 1.5, 1.5, 2.0))

flextable::save_as_docx(ft6, path = here("outputs", "tables", "supplementary", "tableS6_validation_detail.docx"))
ft6
```

