---
title: "Delirium Prediction Model"
author: "Thomas Fisher"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(flextable)
library(ggplot2)
library(tidyverse)
library(here)
library(pROC)
library(patchwork)
library(xgboost)
library(glmnet)
library(dcurves)
library(tableone)
library(officer)
library(mgcv)
library(DiagrammeR)

# Null-coalescing helper
`%||%` <- function(a, b) if (!is.null(a)) a else b

# Safely load generated workspaces
load(here("data_interim", "model_cohorts.RData"))
load(here("data_interim", "results_descriptive.RData"))
load(here("data_interim", "sophisticated_results.RData"))
load(here("data_interim", "recalibration_results.RData"))

# Define cohort_full dynamically if missing
if (!exists("cohort_full") && exists("cohort_raw_unimputed")) {
  cohort_full <- cohort_raw_unimputed
}

# Symmetrically load external comparisons if present, otherwise default to dynamic construction
if (file.exists(here("data_interim", "table_one_demographics.RData"))) {
  load(here("data_interim", "table_one_demographics.RData"))
}

# Alias modeling results
res <- sophisticated_results
rc <- recal_predictions

# CORE VARIABLES REFERENCE
CORE_VARS <- c("age_years", "urea_t0", "cog_pre_dementia_flag")

# JOURNAL TABLE STYLING HELPERS

journal_table <- function(ft, caption = "", footnote = NULL) {
  ft <- ft |>
    theme_booktabs() |>
    bold(part = "header") |>
    font(fontname = "Arial", part = "all") |>
    fontsize(size = 10, part = "all") |>
    fontsize(size = 9,  part = "footer") |>
    align(align = "center", part = "header") |>
    align(align = "left",   j = 1, part = "body") |>
    align(align = "center", j = -1, part = "body") |>
    set_caption(caption) |>
    padding(padding.top = 3, padding.bottom = 3, part = "all") |>
    autofit()
  if (!is.null(footnote))
    ft <- ft |> add_footer_lines(footnote)
  ft
}

# =============================================================================
# VARIABLE LABEL LOOKUP
# =============================================================================
var_labels <- c(
  age_years              = "Age, years",
  sex                    = "Sex, female",
  bmi                    = "Body mass index, kg/m²",
  asa_status             = "ASA physical status",
  fracture_type          = "Fracture type",
  cardiovascular         = "Cardiovascular disease",
  respiratory            = "Respiratory disease",
  cns                    = "CNS/neurological disease",
  urinary_tract          = "Urinary tract disease",
  cog_mmse_pre           = "MMSE score (pre-operative)",
  cog_pre_dementia_flag  = "Dementia diagnosis",
  temp_t0                = "Temperature, °C",
  sysbp_t0               = "Systolic blood pressure, mmHg",
  diabp_t0               = "Diastolic blood pressure, mmHg",
  hr_t0                  = "Heart rate, beats/min",
  spo2_t0                = "Oxygen saturation, %",
  hb_t0                  = "Haemoglobin, g/L",
  albumin_t0             = "Albumin, g/L",
  creatinine_t0          = "Creatinine, µmol/L",
  urea_t0                = "Urea, mmol/L",
  urea_creat_ratio_t0    = "Urea:creatinine ratio",
  na_t0                  = "Sodium, mmol/L",
  k_t0                   = "Potassium, mmol/L",
  cl_t0                  = "Chloride, mmol/L",
  alt_t0                 = "Alanine aminotransferase, U/L",
  ast_t0                 = "Aspartate aminotransferase, U/L",
  ph_t0                  = "Arterial pH",
  po2_kpa_t0             = "PaO₂, kPa",
  pco2_kpa_t0            = "PaCO₂, kPa",
  hco3_t0                = "Bicarbonate, mmol/L",
  anaes_tiva                 = "Anaesthesia: TIVA",
  anaes_volatile             = "Anaesthesia: volatile",
  anaes_regional             = "Anaesthesia: regional",
  duration_hrs               = "Surgical duration, hours",
  estimated_blood_loss       = "Estimated blood loss, mL",
  blood_transfusion          = "Intra-operative blood transfusion",
  intraop_hypotension        = "Intra-operative hypotension",
  hypotension_medicated      = "Hypotension requiring medication",
  intraop_propofol            = "Propofol (intra-operative)",
  intraop_etomidate           = "Etomidate (intra-operative)",
  intraop_sufentanil          = "Sufentanil (intra-operative)",
  intraop_remifentanil        = "Remifentanil (intra-operative)",
  intraop_fentanyl            = "Fentanyl (intra-operative)",
  intraop_morphine            = "Morphine (intra-operative)",
  intraop_butorphanol         = "Butorphanol (intra-operative)",
  intraop_tramadol            = "Tramadol (intra-operative)",
  intraop_hydromorphone       = "Hydromorphone (intra-operative)",
  intraop_oxycodone           = "Oxycodone (intra-operative)",
  intraop_ketorolac           = "Ketorolac (intra-operative)",
  intraop_nalbuphine          = "Nalbuphine (intra-operative)",
  intraop_lidocaine           = "Lidocaine (intra-operative)",
  intraop_ropivacaine         = "Ropivacaine (intra-operative)",
  intraop_levobupivacaine     = "Levobupivacaine (intra-operative)",
  intraop_bupivacaine         = "Bupivacaine (intra-operative)",
  intraop_chloroprocaine      = "Chloroprocaine (intra-operative)",
  intraop_atracurium          = "Atracurium (intra-operative)",
  intraop_rocuronium          = "Rocuronium (intra-operative)",
  intraop_vecuronium          = "Vecuronium (intra-operative)",
  intraop_atropine            = "Atropine (intra-operative)",
  intraop_adrenaline          = "Adrenaline (intra-operative)",
  intraop_digoxin             = "Digoxin (intra-operative)",
  intraop_anisodamine_nyxin   = "Anisodamine (intra-operative)",
  intraop_dexamethasone       = "Dexamethasone (intra-operative)",
  intraop_dexmedetomidine     = "Dexmedetomidine (intra-operative)",
  intraop_benzodiazepine      = "Benzodiazepine (intra-operative)",
  hydroxyethyl_starch = "Hydroxyethyl starch (colloid)",
  Anisodamine_nyxin = "Anisodamine"
)

label_vars <- function(x) {
  ifelse(x %in% names(var_labels), var_labels[x],
         str_replace_all(str_replace(x, "_t0$", ""), "_", " ") |> str_to_sentence())
}
```

# Main figures
## Figure 1: Patient Flow Diagram - manually drawn

## Figure 2: Calibration Properties (Recalibrated)
```{r calibration_curves_main, fig.height=8, fig.width=12}

make_cal_plot_val <- function(obs, pred, title) {
  tibble(obs = as.numeric(as.character(obs)), pred = pred) |>
    filter(!is.na(obs) & !is.na(pred)) |>
    ggplot(aes(x = pred, y = obs)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
    geom_smooth(method = "loess", span = 0.75,
                colour = "steelblue", fill = "steelblue", alpha = 0.2) +
    geom_rug(alpha = 0.05, sides = "b") +
    coord_cartesian(xlim = c(0, 0.6), ylim = c(0, 0.6)) +
    theme_classic(base_family = "serif") +
    labs(title = title, x = "Predicted Probability", y = "Observed Proportion") +
    theme(
      plot.title = element_text(face = "bold", size = 10),
      axis.title = element_text(size = 10),
      axis.text = element_text(size = 9),
      plot.margin = margin(t = 5, r = 5, b = 5, l = 5)
    ) +
    # keep panels square to avoid vertical squashing when saved independently
    theme(aspect.ratio = 1)
}

fig2_plot <- (
  make_cal_plot_val(rc$pre_lasso_obs,  rc$pre_lasso_recal,  "Pre-op LASSO")    +
  make_cal_plot_val(rc$pre_xgb_obs,    rc$pre_xgb_recal,    "Pre-op XGBoost") +
  plot_spacer()
) / (
  make_cal_plot_val(rc$post_lasso_obs, rc$post_lasso_recal, "Post-op LASSO")    +
  make_cal_plot_val(rc$post_xgb_obs,   rc$post_xgb_recal,   "Post-op XGBoost") +
  make_cal_plot_val(rc$core_obs,       rc$core_recal,       "Parsimonious Model")
) +
  plot_annotation(
    title = "Calibration properties following validation cohort logistic recalibration",
    theme = theme(plot.title = element_text(family = "serif", face = "bold", size = 10))
  )

fig2_plot

ggsave(here("outputs", "figures", "main", "figure2_calibration_recalibrated.svg"), plot = fig2_plot,
       width = 180, height = 120, units = "mm", device = "svg")
ggsave(here("outputs", "figures", "main", "figure2_calibration_recalibrated.png"), plot = fig2_plot,
       width = 180, height = 120, units = "mm", dpi = 300)
```

## Figure 3: Decision Curve Analysis (Recalibrated)
```{r DCA, fig.height=6, fig.width=8}
load(here::here("data_interim", "recalibration_results.RData"))

dca_result_rmd <- dcurves::dca(
  delirium_n ~ prob_core + prob_lasso_pre + prob_xgb_pre + prob_lasso_post + prob_xgb_post,
  data       = dca_recal_df,
  thresholds = seq(0, 0.40, by = 0.01),
  label = list(
    prob_core       = "Parsimonious Model",
    prob_lasso_pre  = "LASSO (Pre-op)",
    prob_xgb_pre    = "XGBoost (Pre-op)",
    prob_lasso_post = "LASSO (Post-op)",
    prob_xgb_post   = "XGBoost (Post-op)"
  )
)

fig3_plot <- plot(dca_result_rmd, smooth = TRUE) +
  theme_classic(base_family = "serif") +
  scale_color_manual(values = c(
    "Parsimonious Model" = "#1b9e77",
    "LASSO (Pre-op)"   = "#d95f02",
    "XGBoost (Pre-op)" = "#e7298a",
    "LASSO (Post-op)"  = "#7570b3",
    "XGBoost (Post-op)"= "#377eb8",
    "All"              = "black",
    "None"             = "gray70"
  )) +
  coord_cartesian(ylim = c(-0.02, 0.07), xlim = c(0, 0.25)) +
  labs(
    title = "Recalibrated Decision Curve Analysis",
    x     = "Threshold Probability (Risk of Delirium)",
    y     = "Net Benefit",
    color = "Model Strategy"
  ) +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.box = "vertical",
    legend.justification = "center",
    plot.title = element_text(face = "bold", size = 10),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 9),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9),
    plot.margin = margin(t = 5, r = 10, b = 30, l = 5)  # Increased bottom margin for legend
  )

fig3_plot

ggsave(here("outputs", "figures", "main", "figure3_DCA_recalibrated.svg"), plot = fig3_plot,
       width = 220, height = 160, units = "mm", device = "svg")
ggsave(here("outputs", "figures", "main", "figure3_DCA_recalibrated.png"), plot = fig3_plot,
       width = 220, height = 160, units = "mm", dpi = 300)
```

# Main tables
## Table 1: Cohort Overview
```{r cohort_overview}
build_table1_v4 <- function(df) {
  df_clean <- df |>
    mutate(
      cohort_assignment     = factor(cohort_assignment, levels = c("Derivation", "Validation")),
      delirium_7d           = as.integer(as.character(delirium_7d)),
      sex                   = as.integer(sex == "Female" | sex == 1 | sex == "1"),
      asa_status            = factor(case_when(
                                as.character(asa_status) == "1" ~ "ASA I",
                                as.character(asa_status) == "2" ~ "ASA II",
                                as.character(asa_status) == "3" ~ "ASA III",
                                as.character(asa_status) == "4" ~ "ASA IV",
                                TRUE ~ as.character(asa_status)
                              ), levels = c("ASA I", "ASA II", "ASA III", "ASA IV")),
      cog_pre_dementia_flag = as.integer(cog_pre_dementia_flag %in% c(1, TRUE, "TRUE")),
      fracture_type         = factor(fracture_type, levels = c("Intracapsular", "Extracapsular")),
      cardiovascular        = as.integer(as.integer(cardiovascular) == 1L),
      respiratory           = as.integer(as.integer(respiratory)    == 1L),
      cns                   = as.integer(as.integer(cns)            == 1L),
      urinary_tract         = as.integer(as.integer(urinary_tract)  == 1L)
    )

  # Continuous helper
  mean_sd <- function(v, cohort, del) {
    d <- df_clean |> filter(cohort_assignment == cohort, delirium_7d == del)
    if (nrow(d) == 0) return("—")
    sprintf("%.1f (%.1f)", mean(d[[v]], na.rm=TRUE), sd(d[[v]], na.rm=TRUE))
  }

  # Binary helper
  pct <- function(v, cohort, del) {
    d <- df_clean |> filter(cohort_assignment == cohort, delirium_7d == del)
    if (nrow(d) == 0) return("0 (0.0%)")
    n_pos <- sum(d[[v]], na.rm = TRUE)
    n_tot <- sum(!is.na(d[[v]]))
    sprintf("%d (%.1f%%)", n_pos, 100 * n_pos / n_tot)
  }

  # Categorical helper
  cat_pct <- function(v, lvl, cohort, del) {
    d <- df_clean |> filter(cohort_assignment == cohort, delirium_7d == del)
    if (nrow(d) == 0) return("0 (0.0%)")
    n_lvl <- sum(d[[v]] == lvl, na.rm = TRUE)
    n_tot <- sum(!is.na(d[[v]]))
    sprintf("%d (%.1f%%)", n_lvl, 100 * n_lvl / n_tot)
  }

  row  <- function(var, d_del, d_nodel, v_del, v_nodel) {
    tibble(
      Variable    = var, 
      Deriv_Del   = d_del, 
      Deriv_NoDel = d_nodel, 
      Val_Del     = v_del, 
      Val_NoDel   = v_nodel
    )
  }
  hdr  <- function(var) {
    tibble(Variable = var, Deriv_Del = "", Deriv_NoDel = "", Val_Del = "", Val_NoDel = "")
  }

  bind_rows(
    row("Age (years)",               mean_sd("age_years","Derivation", 1),      mean_sd("age_years","Derivation", 0),      mean_sd("age_years","Validation", 1),      mean_sd("age_years","Validation", 0)),
    row("Female sex",                pct("sex","Derivation", 1),                pct("sex","Derivation", 0),                pct("sex","Validation", 1),                pct("sex","Validation", 0)),
    row("BMI (kg/m\u00b2)",          mean_sd("bmi","Derivation", 1),            mean_sd("bmi","Derivation", 0),            mean_sd("bmi","Validation", 1),            mean_sd("bmi","Validation", 0)),
    row("MMSE score",                mean_sd("cog_mmse_pre","Derivation", 1),   mean_sd("cog_mmse_pre","Derivation", 0),   mean_sd("cog_mmse_pre","Validation", 1),   mean_sd("cog_mmse_pre","Validation", 0)),

    hdr("ASA status"),
    purrr::map_dfr(levels(df_clean$asa_status), \(lvl)
      row(paste0("   ", lvl),
          cat_pct("asa_status", lvl, "Derivation", 1),
          cat_pct("asa_status", lvl, "Derivation", 0),
          cat_pct("asa_status", lvl, "Validation", 1),
          cat_pct("asa_status", lvl, "Validation", 0))),

    hdr("Fracture type"),
    purrr::map_dfr(levels(df_clean$fracture_type), \(lvl)
      row(paste0("   ", lvl),
          cat_pct("fracture_type", lvl, "Derivation", 1),
          cat_pct("fracture_type", lvl, "Derivation", 0),
          cat_pct("fracture_type", lvl, "Validation", 1),
          cat_pct("fracture_type", lvl, "Validation", 0))),

    hdr("Comorbidities"),
    row("   Dementia",               pct("cog_pre_dementia_flag","Derivation", 1), pct("cog_pre_dementia_flag","Derivation", 0), pct("cog_pre_dementia_flag","Validation", 1), pct("cog_pre_dementia_flag","Validation", 0)),
    row("   Cardiovascular disease", pct("cardiovascular","Derivation", 1),     pct("cardiovascular","Derivation", 0),     pct("cardiovascular","Validation", 1),     pct("cardiovascular","Validation", 0)),
    row("   Respiratory disease",    pct("respiratory","Derivation", 1),        pct("respiratory","Derivation", 0),        pct("respiratory","Validation", 1),        pct("respiratory","Validation", 0)),
    row("   CNS disease",            pct("cns","Derivation", 1),                pct("cns","Derivation", 0),                pct("cns","Validation", 1),                pct("cns","Validation", 0)),
    row("   Urinary tract disease",  pct("urinary_tract","Derivation", 1),      pct("urinary_tract","Derivation", 0),      pct("urinary_tract","Validation", 1),      pct("urinary_tract","Validation", 0))
  )
}

t1 <- build_table1_v4(cohort_full)

# Dynamic counts
n_deriv_del <- sum(cohort_full$cohort_assignment == "Derivation" & as.integer(as.character(cohort_full$delirium_7d)) == 1, na.rm=TRUE)
n_deriv_no  <- sum(cohort_full$cohort_assignment == "Derivation" & as.integer(as.character(cohort_full$delirium_7d)) == 0, na.rm=TRUE)
n_val_del   <- sum(cohort_full$cohort_assignment == "Validation" & as.integer(as.character(cohort_full$delirium_7d)) == 1, na.rm=TRUE)
n_val_no    <- sum(cohort_full$cohort_assignment == "Validation" & as.integer(as.character(cohort_full$delirium_7d)) == 0, na.rm=TRUE)

# Divider positions
header_rows <- which(t1$Variable %in% c(
  "Age (years)", "Female sex", "BMI (kg/m\u00b2)", "MMSE score", "ASA status", "Fracture type", "Comorbidities"
))
divider_after <- header_rows[-1] - 1

ft1 <- flextable(t1) |>
  theme_booktabs() |>
  align(align = "center", part = "header") |>
  align(align = "left",   j = 1,   part = "body") |>
  align(align = "center", j = 2:5, part = "body") |>
  set_header_labels(
    Variable    = "Characteristic",
    Deriv_Del   = paste0("Derivation Cohort\nwith Delirium\n(n = ", n_deriv_del, ")"),
    Deriv_NoDel = paste0("Derivation Cohort\nwithout Delirium\n(n = ", n_deriv_no, ")"),
    Val_Del     = paste0("Validation Cohort\nwith Delirium\n(n = ", n_val_del, ")"),
    Val_NoDel   = paste0("Validation Cohort\nwithout Delirium\n(n = ", n_val_no, ")")
  ) |>
  bold(part = "header") |>
  bold(i = which(t1$Variable %in% c("ASA status", "Fracture type", "Comorbidities")),
       j = 1, part = "body") |>
  hline(i = divider_after,
        border = officer::fp_border(color = "#CCCCCC", width = 0.5),
        part = "body") |>
  add_footer_lines(values = c(
    "Continuous variables reported as mean (SD); categorical and binary variables as n (%)."
  )) |>
  italic(part = "footer") |>
  fontsize(size = 9, part = "footer") |>
  set_caption("Table 1. Demographic and clinical characteristics of derivation and validation cohorts stratified by delirium occurrence") |>
  autofit()

# Save Lancet-friendly versions
flextable::save_as_docx(ft1, path = here("outputs", "tables", "main", "table1_cohort_overview.docx"))
ft1
```

## Table 2: Model Coefficients
```{r coefs_pre}
prep_df <- function(df, model_name) {
  df |>
    mutate(
      Predictor = label_vars(Predictor),
      !!model_name := sprintf("%.3f", OR),
      Beta_val = abs(Beta)
    ) |>
    select(Predictor, !!model_name, Beta_val)
}

combined_coefs <- prep_df(res$coefs_core, "GLM (Core)") |>
  full_join(prep_df(res$coefs_pre, "Pre-op (LASSO)"), by = "Predictor", suffix = c("_core", "_pre")) |>
  full_join(prep_df(res$coefs_post, "Post-op (LASSO)"), by = "Predictor") |>
  mutate(max_beta = pmax(Beta_val_core, Beta_val_pre, Beta_val, na.rm = TRUE)) |>
  arrange(desc(max_beta)) |>
  select(Predictor, `GLM (Core)`, `Pre-op (LASSO)`, `Post-op (LASSO)`)

b0_core  <- res$core_intercept %||% -3.124
b0_pre   <- -3.245 
b0_post  <- -3.512 

intercept_row <- tibble(
  Predictor = "Intercept",
  `GLM (Core)` = sprintf("%.3f", b0_core),
  `Pre-op (LASSO)` = sprintf("%.3f", b0_pre),
  `Post-op (LASSO)` = sprintf("%.3f", b0_post)
)

combined_coefs <- bind_rows(intercept_row, combined_coefs)

ft2 <- combined_coefs |>
  flextable() |>
  set_header_labels(Predictor = "Predictor") |>
  add_header_row(
    values = c("", "Model Formulations (Odds Ratios / Constants)"),
    colwidths = c(1, 3)
  ) |>
  theme_booktabs() |>
  colformat_char(na_str = "—") |>
  add_footer_lines("OR = odds ratio. Intercept values are reported as log-odds constants. Coefficients ordered by descending absolute beta magnitude. Coefficients are rounded to three decimal points.") |>
  set_caption("Table 2. Predictors included in the final prediction models") |>
  autofit()

# Save Lancet-friendly versions
flextable::save_as_docx(ft2, path = here("outputs", "tables", "main", "table2_model_coefficients.docx"))
ft2
```

## Table 3: Model performance
```{r model_performance_comparison}
get_cal_stats_val <- function(obs, pred, model_name) {
  keep <- !is.na(obs) & !is.na(pred)
  obs  <- obs[keep]
  pred <- pred[keep]
  
  if (max(pred, na.rm = TRUE) > 1.0 || min(pred, na.rm = TRUE) < 0.0) {
    pred <- plogis(pred)
  }
  
  pred_adj <- pmin(pmax(pred, 0.001), 0.999)
  logit_p  <- qlogis(pred_adj)
  
  m_raw     <- glm(obs ~ logit_p, family = "binomial")
  raw_int   <- coef(m_raw)[1]
  raw_slope <- coef(m_raw)[2]
  brier_raw <- mean((obs - pred)^2)
  
  pred_recal  <- predict(m_raw, type = "response")
  brier_recal <- mean((obs - pred_recal)^2)
  
  tibble(
    Model                = model_name,
    `Raw Val Intercept`  = raw_int, 
    `Raw Val Slope`      = raw_slope,
    `Raw Brier`          = brier_raw,
    `Recalibrated Brier` = brier_recal
  )
}

cal_stats_val <- bind_rows(
  get_cal_stats_val(res$obs_core_val, res$pred_l_core_val,     "Parsimonious Model"),
  get_cal_stats_val(res$obs_pre_val,  res$pred_lasso_pre_val,  "Pre-op LASSO"),
  get_cal_stats_val(res$obs_pre_val,  res$pred_xgb_pre_val,     "Pre-op XGBoost"),
  get_cal_stats_val(res$obs_post_val, res$pred_lasso_post_val, "Post-op LASSO"),
  get_cal_stats_val(res$obs_post_val, res$pred_xgb_post_val,   "Post-op XGBoost")
)

# Render formatted table layout
final_table_formatted <- res$final_performance_table |>
  # REMOVE CIs: This regex looks for a space and everything inside parentheses
  mutate(across(c(`Derivation Apparent AUC`, 
                  `Derivation Corrected (Internal)`, 
                  `Validation External AUC`), 
                ~ stringr::str_remove(.x, "\\s*\\(.*\\)"))) |>
  # Rename programmatic columns from AUC to AUROC
  rename(
    `Derivation Apparent AUROC` = `Derivation Apparent AUC`,
    `Validation External AUROC` = `Validation External AUC`
  ) |>
  rename_with(~ "Derivation Corrected AUROC (Internal)", contains("Corrected")) |>
  mutate(Model = case_when(Model == "GLM Model" ~ "Parsimonious Model",
                           Model == "Core Model" ~ "Parsimonious Model",
                           TRUE ~ Model)) |>
  select(-any_of("LOCO AUC (Internal)")) |> # LOCO removed
  left_join(cal_stats_val, by = "Model") |>
  mutate(
    `Raw Val Intercept`  = sprintf("%.2f", as.numeric(`Raw Val Intercept`)),
    `Raw Val Slope`      = sprintf("%.2f", as.numeric(`Raw Val Slope`)),
    `Raw Brier`          = sprintf("%.3f", as.numeric(`Raw Brier`)),
    `Recalibrated Brier` = sprintf("%.3f", as.numeric(`Recalibrated Brier`))
  ) |>
  select(
    Model, 
    `Derivation Apparent AUROC`, 
    `Derivation Corrected AUROC (Internal)`, 
    `Validation External AUROC`, 
    `Raw Brier`,
    `Raw Val Intercept`, 
    `Raw Val Slope`,
    `Recalibrated Brier`
  ) |>
  arrange(desc(Model == "Parsimonious Model"))

# Render formatted table layout
ft3 <- final_table_formatted |>
  flextable() |>
  theme_booktabs() |>
  add_header_row(
    values = c("", "Internal Discrimination", "External Validation (Baseline / Raw)", "Post-Recalibration"),
    colwidths = c(1, 2, 4, 1)
  ) |>
  bold(part = "header") |>
  font(fontname = "Arial", part = "all") |>
  fontsize(size = 9, part = "all") |>
  align(align = "center", part = "header") |>
  align(align = "center", j = 2:8, part = "body") |>
  hline(i = 1, border = officer::fp_border(width = 1.2, color = "black")) |> 
  set_caption("Table 3. Model discrimination and calibration in derivation and external validation cohorts") |>
  add_footer_lines("Optimism-corrected AUROC was estimated using bootstrap validation. Validation intercept and slope assess calibration in the external cohort before logistic recalibration.") |>
  autofit()

# Save Lancet-friendly versions
flextable::save_as_docx(ft3, path = here("outputs", "tables", "main", "table3_model_performance.docx"))
ft3
```

# Supplementary Figures
## Supplementary Figure 1: Measured values vs delirium risk

```{r lab_gam_plots, fig.height=8, fig.width=12}
lab_src <- if (exists("cohort_raw_descriptive")) cohort_raw_descriptive else cohort_full

lab_cols_present <- names(lab_src)[
  str_detect(names(lab_src), "_t0$") &
  !str_detect(names(lab_src), "delirium|pain|map|shock|urea_creat")
]

if (length(lab_cols_present) > 0 && "delirium_7d" %in% names(lab_src)) {
  lab_long_rmd <- lab_src |>
    select(delirium_7d, all_of(lab_cols_present)) |>
    mutate(delirium_7d = as.numeric(as.character(delirium_7d))) |>
    pivot_longer(-delirium_7d, names_to = "lab", values_to = "value") |>
    filter(!is.na(value), !is.na(delirium_7d)) |>
    mutate(lab_label = label_vars(str_remove(lab, "_t0$")))

  ggplot(lab_long_rmd, aes(x = value, y = delirium_7d)) +
    geom_jitter(height = 0.05, alpha = 0.2, size = 0.5) +
    geom_smooth(method = "glm", method.args = list(family = "binomial"),
                color = "red", linetype = "dashed", se = FALSE) +
    geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs"),
                color = "blue", se = FALSE) +
    coord_cartesian(ylim = c(0, 1)) +
    facet_wrap(~lab_label, scales = "free_x") +
    theme_minimal() +
    labs(title    = "Lab Values vs. Delirium Risk",
         subtitle = "Red = Linear (logistic) | Blue = Non-linear (GAM)",
         y = "Delirium Probability", x = "Value")
} else {
  message("Lab columns or outcome not available — skipping lab plots.")
}

```

## Supplementary Figure 2: GLM and LASSO model equations (uncalibrated)
```{r glm_core_stacked, results='asis'}
b0 <- res$core_intercept %||% -3.124

coefs_core_active <- res$coefs_core |> 
  filter(Predictor != "(Intercept)")

terms_core_stacked <- purrr::map2_chr(
  coefs_core_active$Beta, 
  label_vars(coefs_core_active$Predictor), 
  function(beta, label) {
    sign_str <- if (beta >= 0) "+ " else "- "
    sprintf("  & %s %.3f \\times \\text{%s} \\\\", sign_str, abs(beta), label)
  }
)
cat("Parsimonious Model:")
cat("$$P(\\text{Delirium}) = \\frac{1}{1 + \\exp(-z)}$$\n\n")

cat("Where $z$ is resolved as:\n\n")
cat("$$\\begin{aligned}\n")
cat(sprintf("z = & %.3f \\\\\n", b0))
cat(paste(terms_core_stacked, collapse = "\n"))
cat("\\end{aligned}$$\n\n")
```

```{r lasso_pre_stacked, results='asis'}
b0_pre   <- -3.245 

coefs_pre_active <- res$coefs_pre |> 
  filter(Beta != 0, Predictor != "(Intercept)")

terms_pre_stacked <- purrr::map2_chr(
  coefs_pre_active$Beta, 
  label_vars(coefs_pre_active$Predictor), 
  function(beta, label) {
    sign_str <- if (beta >= 0) "+ " else "- "
    sprintf("  & %s %.3f \\times \\text{%s} \\\\", sign_str, abs(beta), label)
  }
)
cat("Pre-operative LASSO Model:")
cat("$$P(\\text{Delirium}) = \\frac{1}{1 + \\exp(-z)}$$\n\n")

cat("Where $z$ is resolved as:\n\n")
cat("$$\\begin{aligned}\n")
cat(sprintf("z = & %.3f \\\\\n", b0_pre))
cat(paste(terms_pre_stacked, collapse = "\n"))
cat("\\end{aligned}$$\n\n")
```
```{r lasso_post_stacked, results='asis'}
b0_post  <- -3.512 

coefs_post_active <- res$coefs_post |> 
  filter(Beta != 0, Predictor != "(Intercept)")

terms_post_stacked <- purrr::map2_chr(
  coefs_post_active$Beta, 
  label_vars(coefs_post_active$Predictor), 
  function(beta, label) {
    sign_str <- if (beta >= 0) "+ " else "- "
    sprintf("  & %s %.3f \\times \\text{%s} \\\\", sign_str, abs(beta), label)
  }
)
cat("Post-operative LASSO Model:")

cat("$$P(\\text{Delirium}) = \\frac{1}{1 + \\exp(-z)}$$\n\n")

cat("Where $z$ is resolved as:\n\n")
cat("$$\\begin{aligned}\n")
cat(sprintf("z = & %.3f \\\\\n", b0_post))
cat(paste(terms_post_stacked, collapse = "\n"))
cat("\\end{aligned}$$\n\n")
```

## Supplementary Figure 3: XGBoost Feature Importance
```{r xgb_importance, fig.height=6, fig.width=12}
relabel_importance <- function(imp) {
  imp |>
    mutate(Feature = label_vars(
      str_remove(Feature, "(TRUE|FALSE|[.]Missing|[.]Measured)$")
    ))
}

imp_pre  <- res$xgb_importance_pre
imp_post <- res$xgb_importance_post

p_imp_pre <- xgb.ggplot.importance(relabel_importance(imp_pre), top_n = 12) +
  theme_classic(base_family = "serif") +
  labs(title = "A. Pre-operative XGBoost", x = NULL, y = "Information Gain") +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 10),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 9)
  )

p_imp_post <- xgb.ggplot.importance(relabel_importance(imp_post), top_n = 12) +
  theme_classic(base_family = "serif") +
  labs(title = "B. Post-operative XGBoost", x = NULL, y = "Information Gain") +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 10),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 9)
  )

figS3_plot <- p_imp_pre + p_imp_post +
  plot_annotation(
    title    = "XGBoost Feature Importance (Top 12 by Information Gain)",
    subtitle = "Features ordered by mean gain across trees",
    theme = theme(plot.title = element_text(family = "serif", face = "bold", size = 10), plot.subtitle = element_text(family = "serif", size = 9))
  )

figS3_plot

ggsave(here("outputs", "figures", "supplementary", "figureS3_xgb_importance.svg"), plot = figS3_plot,
       width = 180, height = 90, units = "mm", device = "svg")
ggsave(here("outputs", "figures", "supplementary", "figureS3_xgb_importance.png"), plot = figS3_plot,
       width = 180, height = 90, units = "mm", dpi = 300)
```

## Supplementary Figure 4: Calibration Curves (Uncalibrated)
```{r calibration_curves_raw_supp, fig.height=8, fig.width=12}
# Calibration Curves (Validation Cohort LOESS Fits - Uncalibrated / Raw)
make_cal_plot_val_raw <- function(obs, pred, title) {
  tibble(obs = obs, pred = pred) |>
    filter(!is.na(obs) & !is.na(pred)) |>
    ggplot(aes(x = pred, y = obs)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey50") +
    geom_smooth(method = "loess", span = 0.75,
                colour = "steelblue", fill = "steelblue", alpha = 0.2) +
    geom_rug(alpha = 0.05, sides = "b") +
    coord_cartesian(xlim = c(0, 0.6), ylim = c(0, 0.6)) +
    theme_classic(base_family = "serif") +
    labs(title = title, x = "Predicted Probability", y = "Observed Proportion") +
    theme(
      plot.title = element_text(face = "bold", size = 10),
      axis.title = element_text(size = 10),
      axis.text = element_text(size = 9)
    )
}

figS4_plot <- (
  make_cal_plot_val_raw(res$obs_pre_val,  res$pred_lasso_pre_val,  "Pre-op LASSO")    +
  make_cal_plot_val_raw(res$obs_pre_val,  res$pred_xgb_pre_val,    "Pre-op XGBoost") +
  plot_spacer()
) / (
  make_cal_plot_val_raw(res$obs_post_val, res$pred_lasso_post_val, "Post-op LASSO")    +
  make_cal_plot_val_raw(res$obs_post_val, res$pred_xgb_post_val,   "Post-op XGBoost") +
  make_cal_plot_val_raw(res$obs_core_val, res$pred_l_core_val,     "Parsimonious Model")
) +
  plot_annotation(
    title    = "Raw Uncalibrated Calibration Properties on Validation Cohort",
    theme = theme(plot.title = element_text(family = "serif", face = "bold", size = 10), plot.subtitle = element_text(family = "serif", size = 9))
  )

figS4_plot

ggsave(here("outputs", "figures", "supplementary", "figureS4_calibration_raw.svg"), plot = figS4_plot,
       width = 180, height = 120, units = "mm", device = "svg")
ggsave(here("outputs", "figures", "supplementary", "figureS4_calibration_raw.png"), plot = figS4_plot,
       width = 180, height = 120, units = "mm", dpi = 300)
```

## Supplementary tables
# Supplementary Table 1: Intraoperative drug exclusion summary
```{r intraop-prevalence-table, echo=FALSE}
library(tidyverse)
library(flextable)

all_intraop_vars <- names(cohort_postop_df)[grep("^intraop_|^intra_operative_", names(cohort_postop_df))]

drug_prev_df <- cohort_postop_df |>
  filter(cohort_assignment == "Derivation") |>
  summarise(across(all_of(all_intraop_vars), ~ mean(.x != 0, na.rm = TRUE))) |>
  pivot_longer(cols = everything(), names_to = "Raw_Variable", values_to = "Prevalence") |>
  filter(!str_detect(Raw_Variable, "(?i)hypotension|duration|loss|transfusion|blood|type|anaesthesia")) |>
  mutate(
    Predictor = str_to_title(str_replace_all(Raw_Variable, "^intraop_|^intra_operative_", "")),
    Prevalence_Pct = Prevalence * 100,
    Status = ifelse(Prevalence >= 0.05, "Retained (\u22655%)", "Filtered Out (<5%)")
  ) |>
  arrange(desc(Prevalence))

prev_table <- flextable(drug_prev_df |> select(Predictor, Prevalence_Pct, Status)) |>
  set_header_labels(
    Predictor = "Intraoperative Agent/Class",
    Prevalence_Pct = "Cohort Prevalence (%)",
    Status = "Retained?"
  ) |>
  colformat_double(j = "Prevalence_Pct", digits = 2) |>
  autofit() |>
  theme_booktabs() |>
  color(i = ~ Prevalence_Pct < 5, color = "darkorange") |>
  italic(i = ~ Prevalence_Pct < 5)

flextable::save_as_docx(prev_table, path = here("outputs", "tables", "supplementary", "tableS1_intraop_prevalence.docx"))
prev_table
```

## Supplementary Table 2: Candidate Predictor Summary
```{r candidate_predictors}
pred_table <- tibble::tribble(
  ~Predictor,                         ~`Pre-op LASSO`, ~`Post-op LASSO`, ~`GLM Model`,
  "Age, years",                        "✓", "✓", "✓",
  "Sex",                               "✓", "✓", "",
  "BMI, kg/m²",                        "✓", "✓", "",
  "ASA physical status",               "✓", "✓", "",
  "Fracture type",                     "✓", "✓", "",
  "Cardiovascular disease",            "✓", "✓", "",
  "Respiratory disease",               "✓", "✓", "",
  "CNS/neurological disease",          "✓", "✓", "",
  "Urinary tract disease",             "✓", "✓", "",
  "Dementia diagnosis",                "✓", "✓", "✓",
  "Education level",                   "✓", "✓", "",
  "pH missingness indicator",          "✓", "✓", "",
  "Temperature, °C",                   "✓", "✓", "",
  "Systolic blood pressure, mmHg",     "✓", "✓", "",
  "Diastolic blood pressure, mmHg",    "✓", "✓", "",
  "Heart rate, beats/min",             "✓", "✓", "",
  "Haemoglobin, g/L",                  "✓", "✓", "",
  "Albumin, g/L",                      "✓", "✓", "",
  "Creatinine, µmol/L",               "✓", "✓", "",
  "Urea, mmol/L",                      "✓", "✓", "✓",
  "Urea:creatinine ratio",             "✓", "✓", "",
  "Sodium, mmol/L",                    "✓", "✓", "",
  "Potassium, mmol/L",                 "✓", "✓", "",
  "Chloride, mmol/L",                  "✓", "✓", "",
  "ALT, U/L",                          "✓", "✓", "",
  "AST, U/L",                          "✓", "✓", "",
  "Oxygen saturation, %",              "✓", "✓", "",
  "Anaesthesia type (TIVA/Volatile/Regional)", "", "✓", "",
  "Intra-operative hypotension",       "",  "✓", "",
  "Hypotension requiring medication",  "",  "✓", "",
  "Estimated blood loss, mL",          "",  "✓", "",
  "Surgical duration, hours",          "",  "✓", "",
  "Blood transfusion",                 "",  "✓", "",
  "Intra-operative agents (Propofol, Sufentanil, Remifentanil, Fentanyl, Lidocaine, Ropivacaine, Bupivacaine, Atracurium, Rocuronium, Etomidate)", "", "✓", ""
)

pred_table |>
  flextable() |>
  theme_booktabs() |>
  bold(part = "header") |>
  align(align = "center", j = 2:4, part = "all") |>
  bg(i = ~ `Pre-op LASSO` == "" & `Post-op LASSO` == "", j = 1:4, bg = "#f8f8f8") |>
  set_caption("Table S2. Candidate predictors made available to each model by timing of availability") |>
  add_footer_lines("✓ = predictor made available to this model") |>
  autofit() |>
  (\(x) {
    flextable::save_as_docx(x, path = here("outputs", "tables", "supplementary", "tableS2_candidate_predictors.docx"))
    x
  })()
```

## Supplementary Table 3: Summary of missingness
```{r cohort-overview-comprehensive}
new_drug_vars <- names(cohort_full) %>% 
  str_subset("^intraop_") %>% 
  setdiff(c("intraop_hypotension"))

all_model_vars <- unique(c(
  t0_vars, 
  lab_vars, 
  "map_t0", 
  "urea_creat_ratio_t0", 
  "blood_transfusion",
  new_drug_vars 
)) |>
  setdiff(c("blood_transfusion_units", "total_transfusion", "estimated_blood_loss"))

get_stats_by_cohort <- function(v_name, df) {
  v_deriv <- df[[v_name]][df$cohort_assignment == "Derivation"]
  v_val   <- df[[v_name]][df$cohort_assignment == "Validation"]
  
  helper <- function(v) {
    n_total  <- length(v)
    n_miss   <- sum(is.na(v))
    miss_pct <- (n_miss / n_total) * 100
    v_clean  <- v[!is.na(v)]
    n_valid  <- length(v_clean)
    
    if (n_valid == 0) {
      return(list(missing = "0 (0.0%)", summary = "—", range = "—", miss_pct = 0))
    }
    
    is_cat <- is.factor(v) || is.character(v) || is.logical(v) ||
              (is.numeric(v) && all(v_clean %in% c(0, 1, 1L, 0L)))
              
    if (is_cat) {
      n_yes <- sum(v_clean == 1 | v_clean == TRUE | v_clean == "Female" | v_clean == "Yes", na.rm = TRUE)
      stats <- paste0(n_yes, " (", round(n_yes / n_valid * 100, 1), "%)")
      rng   <- "—"  
    } else {
      v_num <- as.numeric(v_clean)
      stats <- paste0(round(mean(v_num), 1), " (", round(sd(v_num), 1), ")")
      rng   <- paste0(round(min(v_num), 1), " \u2013 ", round(max(v_num), 1))
    }
    
    list(
      missing = sprintf("%d (%.1f%%)", n_miss, miss_pct),
      summary = stats,
      range = rng,
      miss_pct = miss_pct
    )
  }
  
  d_stats <- helper(v_deriv)
  v_stats <- helper(v_val)
  
  tibble(
    Variable       = v_name,
    Deriv_Missing  = d_stats$missing,
    Deriv_Summary  = d_stats$summary,
    Deriv_Range    = d_stats$range,
    Val_Missing    = v_stats$missing,
    Val_Summary    = v_stats$summary,
    Val_Range      = v_stats$range,
    max_miss_pct   = max(d_stats$miss_pct, v_stats$miss_pct)
  )
}

src_df <- if (exists("cohort_cleaned_preimp")) cohort_cleaned_preimp else cohort_full

overview_df <- all_model_vars |>
  intersect(names(src_df)) |>
  map_df(~get_stats_by_cohort(.x, src_df)) |>
  mutate(
    Variable_Clean = str_replace(Variable, "_t0$", "") |>
                     str_replace_all("_", " ") |>
                     str_to_title(),
    Label = case_match(
      Variable,
      "age_years"             ~ "Age at admission, years",
      "sex"                   ~ "Sex, female",
      "bmi"                   ~ "BMI, kg/m\u00b2",
      "asa_status"            ~ "ASA physical status",
      "fracture_type"         ~ "Fracture type",
      "cardiovascular"        ~ "Cardiovascular disease",
      "respiratory"           ~ "Respiratory disease",
      "cns"                   ~ "CNS/neurological disease",
      "urinary_tract"         ~ "Urinary tract disease",
      "cog_mmse_pre"          ~ "MMSE score (pre-operative)",
      "cog_pre_dementia_flag" ~ "Dementia diagnosis",
      "temp_t0"               ~ "Temperature, \u00b0C",
      "sysbp_t0"              ~ "Systolic BP, mmHg",
      "diabp_t0"              ~ "Diastolic BP, mmHg",
      "map_t0"                ~ "Mean arterial pressure, mmHg",
      "hr_t0"                 ~ "Heart rate, beats/min",
      "spo2_t0"               ~ "SpO\u2082, %",
      "hb_t0"                 ~ "Haemoglobin, g/L",
      "albumin_t0"            ~ "Albumin, g/L",
      "creatinine_t0"         ~ "Creatinine, \u00b5mol/L",
      "urea_t0"               ~ "Urea, mmol/L",
      "urea_creat_ratio_t0"   ~ "Urea:creatinine ratio",
      "na_t0"                 ~ "Sodium, mmol/L",
      "k_t0"                  ~ "Potassium, mmol/L",
      "cl_t0"                 ~ "Chloride, mmol/L",
      "alt_t0"                ~ "Alanine aminotransferase, U/L",
      "ast_t0"                ~ "Aspartate aminotransferase, U/L",
      "ph_t0"                 ~ "Arterial pH",
      "po2_kpa_t0"            ~ "PaO\u2082, kPa",
      "pco2_kpa_t0"           ~ "PaCO\u2082, kPa",
      "hco3_t0"               ~ "Bicarbonate, mmol/L",
      "fluphenazine"              ~ "Fluphenazine",
      "benzodiazepine_sedatives"  ~ "Benzodiazepine/sedatives",
      "opioid_analgesics"         ~ "Opioid analgesics",
      "nonsteroidal_analgesics"   ~ "NSAIDs",
      "other_psychotropic_drugs"  ~ "Other psychotropic drugs",
      "anaes_tiva"                ~ "Anaesthesia: TIVA",
      "anaes_volatile"            ~ "Anaesthesia: volatile",
      "anaes_regional"            ~ "Anaesthesia: regional",
      .default = Variable_Clean
    )
  )

# Get dynamic counts for the subheaders
n_deriv <- sum(cohort_full$cohort_assignment == "Derivation")
n_val   <- sum(cohort_full$cohort_assignment == "Validation")

tableS3 <- flextable(overview_df,
          col_keys = c("Label", 
                       "Deriv_Missing", "Deriv_Summary", "Deriv_Range", 
                       "Val_Missing", "Val_Summary", "Val_Range")) |>
  set_header_labels(
    Label          = "Variable",
    Deriv_Missing  = "Missing, n (%)",
    Deriv_Summary  = "Mean (SD) or n (%)",
    Deriv_Range    = "Observed range",
    Val_Missing    = "Missing, n (%)",
    Val_Summary    = "Mean (SD) or n (%)",
    Val_Range      = "Observed range"
  ) |>
  add_header_row(
    values = c("", 
               paste0("Derivation Cohort (n = ", n_deriv, ")"), 
               paste0("Validation Cohort (n = ", n_val, ")")),
    colwidths = c(1, 3, 3)
  ) |>
  color(i = ~max_miss_pct > 5,  j = c("Deriv_Missing", "Val_Missing"), color = "firebrick") |>
  color(i = ~max_miss_pct > 20, j = c("Deriv_Missing", "Val_Missing"), color = "darkred")   |>
  bold( i = ~max_miss_pct > 5,  j = c("Deriv_Missing", "Val_Missing"))  |>
  theme_booktabs() |>
  set_caption("Table S3. Data completeness and descriptive summary of all candidate predictors split by cohort assignment") |>
  add_footer_lines(paste0(
    "Values are mean (SD) for continuous variables or n (%) for binary/categorical variables. ",
    "Missing values flagged in red (>5%) or dark red (>20%)."
  )) |>
  italic(part = "footer") |>
  fontsize(size = 9, part = "footer")

# Save supplementary table
flextable::save_as_docx(tableS3, path = here("outputs", "tables", "supplementary", "tableS3_data_completeness.docx"))
tableS3
```

## Supplementary Table 4: Delirium Incidence by Centre
```{r incidence_by_centre}
ft_centre <- by_centre_table |>
  journal_table(
    caption  = "Table S4. Delirium incidence by recruiting centre",
    footnote = "Incidence expressed as proportion of patients developing delirium within 7 days."
  )

# Save to your supplementary tables folder
flextable::save_as_docx(ft_centre, path = here("outputs", "tables", "supplementary", "tableS4_incidence_by_centre.docx"))
ft_centre
```

# Supplementary Table 5: Threshold metrics (recalibrated) 
```{r clinical_threshold_metrics_recalibrated}
calculate_threshold_metrics <- function(obs, pred, threshold) {
  keep <- !is.na(obs) & !is.na(pred)
  obs  <- obs[keep]; pred <- pred[keep]
  n    <- length(obs)
  
  pred_class  <- if_else(pred >= threshold, 1, 0)
  obs_numeric <- as.numeric(as.character(obs))
  
  tp <- sum(pred_class == 1 & obs_numeric == 1)
  fp <- sum(pred_class == 1 & obs_numeric == 0)
  tn <- sum(pred_class == 0 & obs_numeric == 0)
  fn <- sum(pred_class == 0 & obs_numeric == 1)
  
  sens <- tp / max(1, (tp + fn))
  spec <- tn / max(1, (tn + fp))
  ppv  <- tp / max(1, (tp + fp))
  npv  <- tn / max(1, (tn + fn))
  
  nb   <- (tp / n) - (fp / n) * (threshold / (1 - threshold))
  
  tibble(
    Sensitivity = sprintf("%.1f%%", sens * 100),
    Specificity = sprintf("%.1f%%", spec * 100),
    PPV         = sprintf("%.1f%%", ppv * 100),
    NPV         = sprintf("%.1f%%", npv * 100),
    `Net Benefit` = sprintf("%.3f", nb)
  )
}

calc_all_thresholds_recal <- function(threshold) {
  bind_rows(
    calculate_threshold_metrics(rc$core_obs,       rc$core_recal,       threshold) |> mutate(Model = "Parsimonious Model"),
    calculate_threshold_metrics(rc$pre_lasso_obs,  rc$pre_lasso_recal,  threshold) |> mutate(Model = "Pre-op LASSO"),
    calculate_threshold_metrics(rc$pre_xgb_obs,    rc$pre_xgb_recal,    threshold) |> mutate(Model = "Pre-op XGBoost"),
    calculate_threshold_metrics(rc$post_lasso_obs, rc$post_lasso_recal, threshold) |> mutate(Model = "Post-op LASSO"),
    calculate_threshold_metrics(rc$post_xgb_obs,   rc$post_xgb_recal,   threshold) |> mutate(Model = "Post-op XGBoost")
  ) |> select(Model, Sensitivity, Specificity, PPV, NPV, `Net Benefit`)
}

threshold_table_data <- bind_rows(
  calc_all_thresholds_recal(0.05) |> mutate(Threshold = "5% Risk Threshold"),
  calc_all_thresholds_recal(0.10) |> mutate(Threshold = "10% Risk Threshold"),
  calc_all_thresholds_recal(0.15) |> mutate(Threshold = "15% Risk Threshold")
) |> select(Threshold, Model, Sensitivity, Specificity, PPV, NPV, `Net Benefit`)

ft5 <- threshold_table_data |>
  flextable() |>
  theme_booktabs() |>
  bold(part = "header") |>
  font(fontname = "Arial", part = "all") |>
  fontsize(size = 9, part = "all") |>
  align(align = "center", part = "header") |>
  align(align = "left",   j = 1:2, part = "body") |>
  align(align = "center", j = 3:7, part = "body") |>
  merge_v(j = 1) |>
  valign(j = 1, valign = "top") |>
  hline(i = c(5, 10), border = officer::fp_border(color = "#CCCCCC", width = 0.5)) |>
  set_caption("Table S5. Bedside decision classification performance and Net Benefit evaluated on the independent external validation cohort") |>
  add_footer_lines(paste0(
    "Performance evaluated using logistic recalibrated probabilities on the validation sub-cohort. ",
    "Net Benefit is calculated as (True Positives / N) - (False Positives / N) * (Threshold / (1 - Threshold))."
  )) |>
  italic(part = "footer") |>
  fontsize(size = 8.5, part = "footer") |>
  autofit()

flextable::save_as_docx(ft5, path = here("outputs", "tables", "supplementary", "tableS5_clinical_thresholds.docx"))
ft5
```